---
title: "Risks and Odds (v2)"
subtitle: "Risks and Odds Are Not The Same"
author: "Andy Grogan-Kaylor"
date: "today"
format:
  html:
    toc: true
    number-sections: true
    anchor-sections: true
    theme:
      light: yeti
      dark: vapor
    lightbox: true
---

```{r}
#| message: false
#| echo: false

library(plotly)

library(DT)

library(crosstalk)

library(shiny)

```


```{r}
#| message: false
#| echo: false

events <- rep(100, 10)

x_happens <- seq(10, 100, 10)

risk <- x_happens / events

odds <- ((x_happens / events)) / ((100 - x_happens) / events)

odds <- round(odds, 2)

mydata <- data.frame(events, x_happens, risk, odds)

```

# Risk

Let's think about risk: 

$$\text{risk} = P(\text{event happened}) = \frac{\text{number of events}}{\text{number of events} + \text{number of non-events}}$$

# Odds

The odds are the probability that an event happened divided by the probability that it did not happen

$$\text{odds} = \frac{P(\text{event happened})}{P(\text{event didn't happen})}$$

which in turn is equivalent to

$$= \frac{\frac{\text{number of events}}{\text{number of events} + \text{number of non-events}}}{\frac{\text{number of non-events}}{\text{number of events} + \text{number of non-events}}}$$

which incidentally reduces to

$$= \frac{\text{number of events}}{\text{number of non-events}}$$

# Visualize The Risk And Odds

Imagine an *Event X*. This event could be entering a program, exiting a program, or getting a diagnosis of a mental health or physical health condition. 

Click on the table below and/or hover over the graph below to explore various scenarios.

```{r}
#| message: false
#| echo: false

hk <- highlight_key(mydata)

p1 <- plot_ly(hk,
              marker = list(size = 15),
              width = "70%") %>% 
  highlight() %>%
  add_trace(x = ~x_happens, 
            y = ~risk,
            text = ~paste("X Happens", 
                          x_happens, 
                          "times",
                          "\n",
                          "Risk = ", 
                          x_happens, 
                          "/100 = ", 
                          risk),
            hoverinfo = 'text') %>%
  add_trace(x = ~x_happens, 
            y = ~odds, 
            text = ~paste("X Happens", 
                          x_happens, 
                          "times",
                          "\n",
                          "Odds = ", 
                          x_happens, 
                          "/", 
                          100 - x_happens, 
                          " = ", 
                          odds),
            hoverinfo = 'text') %>%
  layout(xaxis = list(title='How many times does X happen?',
                      showspikes = TRUE),
         yaxis = list(title='Risk or Odds',
                      showspikes = TRUE),
         showlegend = FALSE) 

t1 <- DT::datatable(hk, 
                    class="compact",
                    width = "70%",
                    rownames = FALSE,
                    colnames = c('Total Events', 
                                 'How Many Times Does X Happen?', 
                                 'Risk', 
                                 'Odds'),
                    options = list(dom = 't', 
                                   scrollX = T))

# filter_slider("s1", 
#               "How many times does X happen?", 
#               hk, 
#               column = ~x_happens,
#               step = 10)

# bscols(t1, p1)

# t1
# 
# p1

```

## Visualization

```{r}
#| label: fig-risks-odds
#| fig-cap: "Visualization of Risks and Odds"
#| message: false
#| echo: false

p1

```

## Table

```{r}
#| label: tbl-risks-odds
#| tbl-cap: "Table of Risks and Odds"
#| message: false
#| echo: false

t1

```






