---
title: "Logistic Regression Equation"
author: "Andy Grogan-Kaylor"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    highlight: haddock
    toc: yes
  pdf_document: 
    toc: yes
    number_sections: yes
    highlight: haddock
    latex_engine: xelatex
    fig_height: 3
mainfont: Arial
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

# Logistic Regression 

## Equation

Logistic regression--written here with a single independent variable--models the log odds of an outcome as a function of a set of covariates:

$$\ln \Big(\frac{p(outcome)}{1-p(outcome)} \Big) = \beta_0 + \beta_1 x_1$$ 

Here $p(outcome)$ is the probability of the outcome. 

$\frac{p(outcome)}{1-p(outcome)}$ is the *odds* of the outcome.

Hence, $\ln \Big(\frac{p(outcome)}{1-p(outcome)} \Big)$ is the *log odds*.

Logistic regression returns a $\beta$ coefficient for each independent variable $x$.

These $\beta$ coefficients can then be *exponentiated* to obtain *odds ratios*:  $OR = e^{\beta}$

## Rewriting The Equation

We can take the equation:

$$\ln \Big(\frac{p(outcome)}{1-p(outcome)}\Big) = \beta_0 + \beta_1 x_1$$
We exponentiate both sides of the equation:

$$\frac{p(outcome)}{1-p(outcome)} = e^{\beta_0 + \beta_1 x_1}$$

We multiply both sides by the denominator of the fraction that is on the left hand side of the equation:  

$$p(outcome) = e^{\beta_0 + \beta_1 x_1}(1-p(outcome))$$

Then: 

$$p(outcome) = e^{\beta_0 + \beta_1 x_1} - e^{\beta_0 + \beta_1 x_1} * p(outcome)$$
Then:

$$p(outcome) + e^{\beta_0 + \beta_1 x_1} * p(outcome) = e^{\beta_0 + \beta_1 x_1}$$

Then: 

$$(1 + e^{\beta_0 + \beta_1 x_1}) * p(outcome) = e^{\beta_0 + \beta_1 x_1}$$

And, finally: 

$$p(outcome) = \frac{e^{\beta_0 + \beta_1 x_1}}{1 +e^{\beta_0 + \beta_1 x_1}}$$

We sometimes use a shorthand, and say 

$$F(z)=\frac{e^z}{1+e^z}$$

# Graph

We graph a logistic distribution with $\beta_0$ set to 0, and $\beta_1$ set to 1.

```{r}

N <- 100

x <- runif(N, -10, 10)

y <- exp(x)/(1 + exp(x))

library(ggplot2)

library(ggthemes)

ggplot(data = NULL,
       aes(x = x,
           y = y)) +
  geom_point() + 
  labs(title = "Logistic Function",
       y = "Probability") +
  theme_minimal()

```

