---
title: "Grading Rubric"
subtitle: "SW 672"
author: "Andy Grogan-Kaylor"
institute: "University of Michigan"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: animate.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}

@import url('https://fonts.googleapis.com/css2?family=Montserrat&display=swap');
.title-slide { 
  color: white;
  background-color: black; 
}
.title-slide h1 {
    color: white;
    font-weight: bold;
}
pre {
  white-space: pre-wrap;
}
h1, h2, h3 {
  font-family: 'Montserrat', sans-serif;
}
body { 
    font-family: 'Montserrat', sans-serif;
}
blockquote {
  background: #FFFFFF;
  color: #000;
  border-left: 5px solid #EC1F27;
  margin: 1.5em 10px;
  padding: 0.5em 10px;
}
.author, .date {
  font-family: 'Montserrat', sans-serif;
}
.emphasized {
  background-color: black;
  color: white;
  font-size: 200%;
  text-align: center;
  text-shadow: 0 0 20px #333;
}

```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

xaringanExtra::use_tile_view()

# xaringanExtra::use_slide_tone()

```

# Navigation

* Use the <span style="font-size:100px">&#8678;</span> and <span style="font-size:100px">&#8680;</span> keys to move through the presentation.
* Press *o* for *panel overview*.

---
# The Basics

```{r, fig.height=2, fig.width=4}

library(DiagrammeR)

grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, 
  fontsize = 10,
  rankdir = LR]

  # several 'node' statements
  node [shape = circle,
        fontname = Helvetica]
  A; B

  # several 'edge' statements
  A->B
}
")

```

--

```{r, fig.height=3, fig.width=4, fig.retina=3}

N <- 100

A <- rnorm(N, 10, 1)

e <- rnorm(N, 0, 1)

B <- A + e

plot(A, B, 
     main = "A Is Associated With B", 
     col = rainbow(9),
     pch = 19)

# mean of B by A > mean(A)

aggdata <- aggregate(B, 
                     by = list(A >= mean(A)), 
                     FUN=mean)

barplot(aggdata$x,
        main = "A Is Associated With B", 
        ylab = "Mean of B",
        col = c("red", "black"),
        names.arg = c("A Is Low",
                      "A Is High"),
        cex.names = 1.0)

```


---
# A Conceptual Focus 
## (Grading Rubric 1)

* Disparities or Injustices: How does membership in **A** affect the chances of outcome **B**?

--

* Effect of a Policy or Intervention: Do those who receive treatment or policy **A** have a different outcome **B**?

---
class: animated, slideInRight

# Think Carefully About Design Elements 
## (Grading Rubric 2)

1. Geometry: bars; lines; points; etc.
2. Use of color
3. Well labeled axes
4. Good informative title

```{r}

x <- round(rnorm(100, 100, 10), digits = 0)

```

```{r, eval=TRUE, fig.height=3, fig.retina=2}

library(ggplot2)

library(ggbeeswarm)

ggplot(data = NULL,
       aes(x = x,
           fill = factor(x))) +
  geom_histogram() +
  geom_label(aes(x = 120, 
                 y = 10, 
                 label = "possibly \ninformative \nkey takeaway"),
             size = 2, 
             fill = "white",
             color = "red") +
  ylab("possibly another \ndimension of the data") +
  xlab("one dimension of the data") +
  labs(title = "Informative Title",
       subtitle = "Maybe The Title Is Even Your Main Takeaway",
       caption = "Every extra 'dimension' (x axis, y axis, color, shape) 
       should convey information.
       \n Words should be spelled out.") + 
  theme_minimal() +
  theme(legend.position = "none")

```


---
# Wrap Your Visual In A Narrative 
## (Grading Rubric 3)

```{r, fig.height=2, fig.width=4}

grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, 
  fontsize = 10,
  rankdir = LR]

  # several 'node' statements
  node [shape = circle,
        fontname = Helvetica]
  A; B

  # several 'edge' statements
  A->B
}
")

```

* Why is *A* important? Is *A* prevalent? Widespread? Serious?

--

and / or ...

* Why is *B* important? Is *B* prevalent? Widespread? Serious?

---

* If *A* is connected to *B*, this means: __________

--

and / or ...

* If *A* is not connected to *B*, this means: __________

--

* These findings suggest the following intervention / treatment / policy / advocacy effort: __________

---
class: center, emphasized

We have a saying in Chile: criticizing a musician is easy, but it is more difficult "when you have a guitar in your hand."

-César Hidalgo (L’Université Fédérale Toulouse Midi-Pyrénées)

---
class: animated, slideInRight

# The Difficult Stuff 

(Which I Am Still Figuring Out Best Practices For)

> NOT REQUIRED

* More thoughtful design: Cedric Sherer has an [amazing article on improving the design of a ggplot](https://cedricscherer.netlify.app/2019/05/17/the-evolution-of-a-ggplot-ep.-1/).
* Causality (Does A *cause* B?) [https://agrogan.shinyapps.io/causality/](https://agrogan.shinyapps.io/causality/)
* Moderation (Does C *affect the relationship* between A and B?)
* Simpson's Paradox (When we think about C, does that *reverse* the observed association between A and B?): [https://agrogan.shinyapps.io/simpsonsparadox/](https://agrogan.shinyapps.io/simpsonsparadox/)
* Bayes Theorem (Does thinking about the prior information affect the conclusions we draw?): [https://agrogan.shinyapps.io/Thinking-Through-Bayes/](https://agrogan.shinyapps.io/Thinking-Through-Bayes/)


---
class: center, emphasized

# Questions? 






