---
title: "Times, Events, and Censoring"
author: "Andy Grogan-Kaylor"
date: "`r format(Sys.Date())`"
output:
  html_document: 
    code_folding: hide
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    css: ../UM.css
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

```

# Simulate Data

```{r}

N <- 10 # sample size

T <- 10 # number of timepoints

id <- rep(seq(1, N), each = T) # id's 1 to N

t <- rep(seq(1, T), N) # timepoints 1 to T

# random event times

# uniform event times

event_time <- rep(round(runif(N, 3, T), 
                        digits = 0), 
                  each = T)

event <- t >= event_time # event has occurred t >= event_time

event <- factor(event, 
                levels = c(FALSE, TRUE), 
                labels = c("No Event", "Event"))

# arbitrarily censored after 7

censored <- event_time > 7 

censored <- factor(censored, 
                   levels = c(FALSE, TRUE), 
                   labels = c("Not Censored", "Censored"))

# tibble

mydata <- tibble::tibble(id, t, event_time, event, censored)


```

# Animated Graph

```{r}

library(ggplot2)

library(plotly)

```

Think about a hypothetical event: e.g. *birth*, *death*, *marriage* or *commitment* to a partner, *entering a program*, *leaving a program*.

Individuals in the animation below who have not yet experienced an event are indicated by a $\bullet$. When an event occurs for an individual, the symbol changes to a $\boldsymbol \times$.

In this simulation, we imagine that the study period ends after time 7, so observations for which the event occurs after time 7 are considered to be *censored*. Censored observations are <span style="color: #0000FF">blue</span>, and non-censored observations are <span style="color: #FF0000">red</span>.

```{r}

pal <- c("red", "blue") # color palette

p2 <- plot_ly(data = mydata, # use mydata
              x = ~t, # x is t (time)
              y = ~id, # y is id
              frame = ~t, # frames based on t (time)
              text = ~paste("Event:" ,
                            event,
                            "<br>Censored:",
                            censored),
              type = 'scatter',
              mode = 'marker',
              color = ~censored, # color is censored (yes / no)
              colors = pal, # use color palette
              symbol = ~as.numeric(event), # symbol is event (occurred / not occurred)
              symbols = c('circle', # event not occurred
                          'x'), # event
              marker = list(size = 10)) %>% # marker size
  layout(title = 'Hypothetical Timing of Events \nCensored at Time 7',
         shapes = list(type='line', # censoring line
                       x0 = 7, 
                       x1 = 7, 
                       y0 = 0, 
                       y1 = 10, 
                       line=list(dash='dot', width=1))) %>%
  animation_opts(3000) 

p2 # replay

```

# What These Data Look Like

Notice how every row in this particular data set is a *person timepoint*, **not** simply a *person*.  Every *person* in this data has *multiple rows*.

```{r}

library(DT)

datatable(mydata)

```



