---
title: "ggplot flipbook"
subtitle: "Use <span style='font-size:50px'>&#8678;</span> and <span style='font-size:50px'>&#8680;</span> to navigate."
author: "Andy Grogan-Kaylor"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu, animate.css]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}

@import url('https://fonts.googleapis.com/css2?family=Montserrat&display=swap');
.title-slide { 
  color: #ffcb05;
  background-color: #00274C; 
}
.title-slide h1 {
    color: #ffcb05;
}
pre {
  white-space: pre-wrap;
}
h1, h2, h3 {
  font-family: 'Montserrat', sans-serif;
}
body { 
    font-family: 'Montserrat', sans-serif;
}
.author, .date {
  font-family: 'Montserrat', sans-serif;
}

```

```{r setup, include = F}

# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming

knitr::opts_chunk$set(fig.width = 6, 
                      message = FALSE, 
                      warning = FALSE, 
                      comment = "", 
                      cache = FALSE, 
                      fig.retina = 3)

library(flipbookr)

library(tidyverse)

library(xaringanExtra)

xaringanExtra::use_tile_view()

xaringanExtra::use_slide_tone()

xaringanExtra::use_share_again()


```

```{r, echo = FALSE}

N <- 100

predictor <- rnorm(N, 100, 10)

e <- rnorm(N, 0, 5)

outcome <- predictor + e

group <- c(rep("A", 30),
           rep("B", 70))

mydata <- data.frame(predictor, outcome, group)

```

# How To Navigate This Presentation

* Use the <span style="font-size:100px">&#8678;</span> and <span style="font-size:100px">&#8680;</span> keys to move through the presentation.
* Press *o* for *panel overview*.
* This presentation plays a *tone* when each new item appears. Turn the volume down if you find these tones annoying.

Based upon [2 Page ggplot](https://agrogan1.github.io/R/two-page-ggplot2/two-page-ggplot2.pdf)

---
class: animated, slideInRight
# The Essential Idea Of ggplot2 Is Simple 

There are 3 essential elements to any ggplot call:

1. An *aesthetic* that tells ggplot which variables are being mapped to the *x axis*, *y axis*, (and often other attributes of the graph, such as the *color fill*).  Intuitively, the aesthetic can be thought of as **what you are graphing**.
2. A *geom* or *geometry* that tells ggplot about the basic structure of the graph.  Intuitively, the geom can be thought of as **how you are graphing it**.
3. Other options, such as a *graph title*, *axis labels* and *overall theme* for the graph.

---
class: animated, slideInRight
# Get Started

```{r graph0, include = FALSE}

library(ggplot2) # beautiful graphs

library(ggthemes) # nice themes 

```

---

```{r, code = knitr::knit_code$get("graph0"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph0", break_type = "auto", widths = c(4,2))`

---
class: animated, slideInRight

# One Continuous Variable

```{r graph1, include = FALSE}

ggplot(mydata, # data
       aes(x = outcome)) + # variable
  geom_histogram() #BREAK2

```

---

```{r, code = knitr::knit_code$get("graph1"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph1", break_type = "auto", widths = c(4,2))`

---
class: animated, slideInRight

# Add Color

```{r graph1colorful, include = FALSE}

ggplot(mydata, # data
       aes(x = outcome)) + # variable
  geom_histogram(fill = "red") #BREAK2

```

---

```{r, code = knitr::knit_code$get("graph1colorful"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph1colorful", break_type = "auto", widths = c(4,2))`

---
class: animated, slideInRight

# An Alternative Geometry

```{r graph1A, include = FALSE}

ggplot(mydata, # data
       aes(x = outcome)) + # variable
  geom_density(fill = "blue") #BREAK2

```

---

```{r, code = knitr::knit_code$get("graph1A"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph1A", break_type = "auto", widths = c(4,2))`

---
class: animated, slideInRight

# One Categorical Variable

```{r graph2, include=FALSE}

ggplot(mydata, # data
       aes(
         x = group, 
         fill = group #BREAK2
       )) + 
  geom_bar() # geometry  

```

---

```{r, code = knitr::knit_code$get("graph2"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph2", break_type = "non_seq", widths = c(4,2))`

---
class: animated, slideInRight

# Continuous by Continuous

```{r graph3, include=FALSE}

ggplot(mydata, # data
       aes(
         x = predictor, # variables
         y = outcome, 
         color = group #BREAK2
       )) + 
  geom_point() # geometry

```

---

```{r, code = knitr::knit_code$get("graph3"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph3", break_type = "non_seq", widths = c(4,2))`

---
class: animated, slideInRight

# Build A Graph With Many Parts

```{r graph4, include=FALSE}

ggplot(mydata, # data
       aes(x = predictor, # variables
           y = outcome,
           color = group)) + 
  geom_point() + # 1st geometry
  geom_smooth() + # 2nd geometry
  labs(title = "My Title", caption = "Colors Inspired by UNICEF", x = "Predictor", y = "Outcome") + # labels
  scale_colour_manual(name = "Group", values = c("#FF0099", "#0099FF"),  labels = c("group 1", "group 2")) + # manual colors 
  theme_minimal() + # theme
  theme(plot.title = element_text(size = 20, color = "#0099FF"), plot.caption = element_text(size=rel(1.25))) # add to theme

```

---

```{r, code = knitr::knit_code$get("graph4"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph4", break_type = "auto", widths = c(4,2))`

---
class: animated, slideInRight

# References

Reynolds, E. (2020). _flipbookr: Parses code, creates
partial code builds, delivers code movie_. R package
version 0.1.0.

Wickham, H. (2016). _ggplot2: Elegant Graphics for Data
Analysis_. Springer-Verlag New York. ISBN:
978-3-319-24277-4. URL:
[https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org).


