---
title: "ggplot flipbook"
subtitle: "Based Upon *2 Page ggplot*"
author: "Andy Grogan-Kaylor"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}

.title-slide { 
  color: #ffcb05;
  background-color: #00274C; 
}
.title-slide h1 {
    color: #ffcb05;
}
pre {
  white-space: pre-wrap;
}

```

```{r setup, include = F}

# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming

knitr::opts_chunk$set(fig.width = 6, 
                      message = FALSE, 
                      warning = FALSE, 
                      comment = "", 
                      cache = FALSE, 
                      fig.retina = 3)

library(flipbookr)

library(tidyverse)

```

```{r, load_refs, include=FALSE, cache=FALSE}

library(RefManageR)

BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)

myBib <- ReadBib("./ggplot-flipbook.bib", check = FALSE)

```

```{r, echo = FALSE}

N <- 100

predictor <- rnorm(N, 100, 10)

e <- rnorm(N, 0, 5)

outcome <- predictor + e

group <- c(rep("A", 30),
           rep("B", 70))

mydata <- data.frame(predictor, outcome, group)

```

---
# Get Started

Use the <span>&#8678;</span> and <span>&#8680;</span> keys to move through the presentation.

Based upon [2 Page ggplot](https://agrogan1.github.io/R/two-page-ggplot2/two-page-ggplot2.pdf)

```{r graph0, include = FALSE}

library(ggplot2) # beautiful graphs

library(ggthemes) # nice themes 

```

---

```{r, code = knitr::knit_code$get("graph0"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph0", break_type = "auto")`

---
# One Continuous Variable

```{r graph1, include = FALSE}

ggplot(mydata, # data
       aes(x = outcome)) + # variable
  geom_histogram() # geometry

```

---

```{r, code = knitr::knit_code$get("graph1"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph1", break_type = "auto")`

---
# One Categorical Variable

```{r graph2, include=FALSE}

ggplot(mydata, # data 
       aes(x = group,
           fill = group)) + # variable
  geom_bar() # geometry

# notice how the use of 
# `fill` in the aesthetic
# generates a legend

```

---

```{r, code = knitr::knit_code$get("graph2"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph2", break_type = "auto")`

---
# Continuous by Continuous

```{r graph3, include=FALSE}

ggplot(mydata, # data
       aes(x = predictor, # variables
           y = outcome, 
           color = group)) + 
  geom_point() + # 1st geometry
  geom_smooth() # 2nd geometry

# notice how the use of 
# `color` in the aesthetic 
# generates a legend

```

---

```{r, code = knitr::knit_code$get("graph3"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph3", break_type = "auto")`

---
# Add Some Options

```{r graph4, include=FALSE}

ggplot(mydata, # data
       aes(x = predictor, # variables
           y = outcome,
           color = group)) + 
  geom_point() + # 1st geometry
  geom_smooth() + # 2nd geometry
  labs(title = "My Title", # title
       x = "Predictor", # x title
       y = "Outcome") + # y title
  # manual colors
  scale_colour_manual(name = "Group", 
                      values = 
                        c("#FF0099", 
                          "#0099FF"),  
                      labels = 
                        c("group1", 
                          "group2")) +
  theme_minimal() + # theme
  # add to theme
  theme(plot.title = 
          element_text(size = 20, 
                       color = "#0099FF")) 

```

---

```{r, code = knitr::knit_code$get("graph4"), eval = FALSE, echo = FALSE}
```

`r chunk_reveal("graph4", break_type = "auto")`

---
# References

Reynolds, E. (2020). _flipbookr: Parses code, creates
partial code builds, delivers code movie_. R package
version 0.1.0.

Wickham, H. (2016). _ggplot2: Elegant Graphics for Data
Analysis_. Springer-Verlag New York. ISBN:
978-3-319-24277-4. URL:
[https://ggplot2.tidyverse.org](https://ggplot2.tidyverse.org).

```{r refs, echo=FALSE, eval=FALSE, results="asis"}

# Cite(myBib)
# PrintBibliography(myBib)

```
